---
title: "Compensation Analysis Report"
author: "Compensation Analytics"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    number-sections: true
    code-fold: true
    self-contained: true
params:
  specialty: "All"
  start_date: !r Sys.Date() - 365
  end_date: !r Sys.Date()
  report_type: "Executive Summary"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(dplyr)
library(ggplot2)
library(plotly)
library(knitr)

# Source analytics functions
source("../R/analytics.R")
source("../R/database.R")

# Parameters
specialty <- params$specialty
start_date <- params$start_date
end_date <- params$end_date
report_type <- params$report_type

# Load data
data <- get_compensation_summary()

# Filter data based on parameters
if (specialty != "All") {
  data <- data %>% filter(specialty == !!specialty)
}
```

# Executive Summary

This report provides compensation analysis for the period from `r format(start_date, "%B %d, %Y")` to `r format(end_date, "%B %d, %Y")`.

```{r metrics}
metrics <- calculate_metrics(data)

summary_stats <- data.frame(
  Metric = c(
    "Total Physicians",
    "Over Target",
    "Under Target",
    "% Over Target",
    "Average Variance %",
    "Total Compensation",
    "Average Compensation"
  ),
  Value = c(
    format(metrics$total_physicians),
    format(metrics$over_target),
    format(metrics$under_target),
    sprintf("%.1f%%", metrics$pct_over_target),
    sprintf("%.2f%%", metrics$avg_variance_pct),
    sprintf("$%.0f", metrics$total_compensation),
    sprintf("$%.0f", metrics$avg_compensation)
  )
)

knitr::kable(summary_stats, caption = "Key Metrics")
```

# Detailed Analysis

## Compensation by Specialty

```{r variance_by_specialty}
specialty_variance <- variance_by_specialty(data)

p <- ggplot(specialty_variance, aes(x = reorder(specialty, -avg_variance_pct), 
                                     y = avg_variance_pct, 
                                     fill = avg_variance_pct > 0)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("red", "green")) +
  labs(
    title = "Average Compensation Variance by Specialty",
    x = "Specialty",
    y = "Variance %"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none")

ggplotly(p)
```

## Variance Table

```{r variance_table}
specialty_variance %>%
  mutate(
    avg_target = sprintf("$%.0f", avg_target),
    avg_actual = sprintf("$%.0f", avg_actual),
    avg_variance = sprintf("$%.0f", avg_variance),
    avg_variance_pct = sprintf("%.2f%%", avg_variance_pct)
  ) %>%
  knitr::kable(caption = "Compensation Analysis Summary")
```

## Outlier Analysis

```{r outliers}
outlier_data <- identify_outliers(data, threshold = 2)
outliers <- outlier_data %>% filter(is_outlier)

if (nrow(outliers) > 0) {
  cat("### Physicians with Significant Variance\n\n")
  outliers %>%
    select(physician_name, specialty, target_comp, actual_comp, variance_pct) %>%
    mutate(
      target_comp = sprintf("$%.0f", target_comp),
      actual_comp = sprintf("$%.0f", actual_comp),
      variance_pct = sprintf("%.2f%%", variance_pct)
    ) %>%
    knitr::kable()
} else {
  cat("No significant outliers detected in this period.\n")
}
```

## Summary Statistics

```{r summary_stats}
stats <- get_summary_stats(data)

summary_table <- data.frame(
  Statistic = c("Minimum", "Q1", "Median", "Mean", "Q3", "Maximum", "Std Dev"),
  Value = c(
    sprintf("$%.0f", stats$min_comp),
    sprintf("$%.0f", stats$q1_comp),
    sprintf("$%.0f", stats$median_comp),
    sprintf("$%.0f", stats$mean_comp),
    sprintf("$%.0f", stats$q3_comp),
    sprintf("$%.0f", stats$max_comp),
    sprintf("$%.0f", stats$sd_comp)
  )
)

knitr::kable(summary_table, caption = "Compensation Distribution Statistics")
```

# Trends

```{r trends}
if ("date" %in% names(data)) {
  trends <- analyze_trends(data)
  
  if (!is.null(trends)) {
    p <- ggplot(trends, aes(x = date, y = avg_comp)) +
      geom_line() +
      geom_point() +
      labs(
        title = "Compensation Trend",
        x = "Date",
        y = "Average Compensation ($)"
      ) +
      theme_minimal()
    
    ggplotly(p)
  }
} else {
  cat("Trend data not available.\n")
}
```

# Recommendations

Based on this analysis, consider the following:

1. **Variance Monitoring**: Continue monitoring specialties with high variance percentages
2. **Outlier Review**: Review outlier cases for additional context and fairness validation
3. **Benchmarking**: Compare results against industry benchmarks for similar specialties
4. **Follow-up**: Schedule reviews with departments showing significant deviations

---

*Report generated: `r format(Sys.time(), "%B %d, %Y at %H:%M:%S")`*
*Specialty Filter: `r specialty`*
*Report Type: `r report_type`*
